Data Munging

### 2021-01-21 scripts/CDF_Estimates_Jb.Rmd

Revised: December 12, 2019 to calculate estimates using the SRS variance
         estimator in addition to the local mean variance estimator

@@

@@

@@

# Print summary information for the response variables
cat("\n\nSummary Information for Response Variables\n")
temp <- match(c("CH4_SAT_RATIO", "CO2_SAT_RATIO", "DELTA_13_CH4_C",
   "DELTA_13_CO2_C", "DISSOLVED_CH4", "DISSOLVED_CO2", "DISSOLVED_N2O",
    "N2O_SAT_RATIO", "F_N2O", "F_N2O_D"), names(cdf))
for(c in temp) {
   cat("\n")
   print(describe(cdf[, c], descript=names(cdf)[c]))
}

@@

# Create the subpop data frame, which defines populations and subpopulations for
# which estimates are desired
subpop <- data.frame(siteID=cdf$SITE_ID,
                     National=rep("National", nr),
                     LAKE_ORGN=cdf$LAKE_ORGN,
                     AREA_CAT6=cdf$AREA_CAT6,
                     #  > 1 hour with all variables and subpopulations
                     # AG_ECO3_NM=cdf$AG_ECO3_NM,
                      AG_ECO9_NM=cdf$AG_ECO9_NM
                     # EPA_REG=cdf$EPA_REG,
                     # MAJ_BAS_NM=cdf$MAJ_BAS_NM,
                     # HUC2_NM=cdf$HUC2_NM,
                     # OWN_NARS=cdf$OWN_NARS,
                     # FS_EW=cdf$FS_EW
)

@@

# Create the data.cont data frame, which specifies the variables to use in the
# analysis.  > 1 hour with all variables.
data.cont <- data.frame(siteID=cdf$SITE_ID,
                        # CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
                        # CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
                        # DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
                        # DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
                        # DISSOLVED_CH4=cdf$DISSOLVED_CH4,
                        # DISSOLVED_CO2=cdf$DISSOLVED_CO2,
                        # DISSOLVED_N2O=cdf$DISSOLVED_N2O,
                        # N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
                        F_N2O = cdf$F_N2O) # diffusive N2O nmol N2O m-2 day-1

@@

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # < 1 hour with all variables. 1 minute with F_N2O

@@

# Write percentile estimates as a csv file
write.csv(CDF_Estimates$Pct, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

@@

#
# Calculate CDF estimates using the SRS variance estimator
#

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
   vartype = "SRS")

# Write CDF estimates as a comma-separated value (csv) file
eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

# Create a PDF file containing plots of the CDF estimates
eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

>>>>

Revision history:
December 12, 2019 to calculate estimates using the SRS variance estimator in addition to the local mean variance estimator.

@@

December 3, 2020 added extent estimate of N~2~O source/sink

@@

# Setup file path
localPath <- Sys.getenv("USERPROFILE")

@@

# Print summary information for the response variables
cat("\n\nSummary Information for Response Variables\n")
temp <- match(c("CH4_SAT_RATIO", "DISSOLVED_CH4", "DELTA_13_CH4_C", "E_CH4_UMOL_D", "F_CH4_KM_D",
                "CO2_SAT_RATIO", "DISSOLVED_CO2", "DELTA_13_CO2_C", "E_CO2_UMOL_D", "F_CO2_KM_D",
                "N2O_SAT_RATIO", "DISSOLVED_N2O", "E_N2O_NMOL_D", "F_N2O_M_D"), names(cdf))
for(c in temp) {
  cat("\n")
  print(describe(cdf[, c], descript=names(cdf)[c]))
}

@@

# Create the subpop data frame, which defines populations and subpopulations for
# which estimates are desired
subpop <- data.frame(siteID=cdf$SITE_ID,
                     National=rep("National", nr),
                     LAKE_ORGN=cdf$LAKE_ORGN,
                     AREA_CAT6=cdf$AREA_CAT6,
                     #AG_ECO3_NM=cdf$AG_ECO3_NM,
                     AG_ECO9_NM=cdf$AG_ECO9_NM
                     #EPA_REG=cdf$EPA_REG,
                     #MAJ_BAS_NM=cdf$MAJ_BAS_NM,
                     #HUC2_NM=cdf$HUC2_NM,
                     #OWN_NARS=cdf$OWN_NARS,
                     #FS_EW=cdf$FS_EW
                     )

@@

# Create the data.cont data frame, which specifies the variables to use in the
# analysis.  30 minutes on VM.
data.cont <- data.frame(siteID=cdf$SITE_ID,
                        CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
                        CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
                        N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
                        DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
                        DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
                        DISSOLVED_CH4=cdf$DISSOLVED_CH4,
                        DISSOLVED_CO2=cdf$DISSOLVED_CO2,
                        DISSOLVED_N2O=cdf$DISSOLVED_N2O,
                        E_CH4_UMOL_D = cdf$E_CH4_UMOL_D, # umol CH4 m-2 day-1
                        F_CH4_KM_D = cdf$F_CH4_KM_D, # kmol CH4 day-1
                        E_CO2_UMOL_D = cdf$E_CO2_UMOL_D, # umol CO2 m-2 day-1
                        F_CO2_KM_D = cdf$F_CO2_KM_D, # kmol CO2 day-1
                        E_N2O_NMOL_D = cdf$E_N2O_NMOL_D, # nmol N2O m-2 day-1
                        F_N2O_M_D = cdf$F_N2O_M_D) # mol N2O day-1

@@

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
time.initial <- Sys.time()
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # 30 minutes on VM
Sys.time() - time.initial
```

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cont.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cont.analysis.\n")
}

# Write percentile estimates as a csv file
write.csv(CDF_Estimates$Pct, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Percentile_Estimates_Local_Mean_JB",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

# Calculate extent estimate of N2O source/sink
```{r}
# Can use same sites, subpop, and design frames from above. 
# Add N2O source/sink column to dg
cdf$N2O_SRC_SNK = with(cdf, ifelse(N2O_SAT_RATIO > 1, "source", "sink"))
data.cat <-  data.frame(siteID=cdf$SITE_ID,
                        N2O_SRC_SNK = cdf$N2O_SRC_SNK)

Extent_Estimate <- cat.analysis(sites, subpop, design, data.cat)

print(Extent_Estimate)
```

# # Write extent estimates as a comma-separated value (csv) file
write.csv(Extent_Estimate, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Extent_Estimates",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

# Calculate CDF estimates using the SRS variance estimator
#

# Calculate the estimates
# cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
# if(exists("warn.df")) rm("warn.df")
# CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
#    vartype = "SRS")

@@

# # Check for warning messages and print them if any exist
# if(exists("warn.df")) {
#    cat("\nWarning messages generated during the call to cat.analysis:\n")
#    warnprnt()
# } else {
#    cat("\nNo warning messages were generated during the call to cat.analysis.\n")
# }
# 
# # Write CDF estimates as a comma-separated value (csv) file
# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
# 
# # Create a PDF file containing plots of the CDF estimates
# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))
# 
# # Write percentile estimates as a csv file
# eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

@@

### 2021-01-21 scripts/dataMunge.Rmd

The dissolved gas data were curated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2021-01-13.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

@@

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2021-01-13.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  filter(!grepl("AK", site.id)) # omit Alaska sites
```

@@

k~600~ is a standardized value that must be corrected for the differences in diffusivity among gases and water temperature at the sampling sites.  These corrections will be implemented at a later date and emission rates will be calculated using k~600~ for now.

@@

# although data I provided Tom (nla17gasAggregatedForPopEst2019-10-23.txt) included data from both visits,
# data used for population estimates included only visit one.  I don't know rationale for this, but will
# do the same with the revised file here.  Also need to strip out air samples.
unique(cdf$VISIT_NO) # 1
dg.1 <- dg %>% filter(visit.no == 1, sample.source == "DG")

# even after filtering, dg has 85 more observations than cdf
dim(cdf) # 1005, 81
dim(dg.1) # 1090, 306

@@

dim(cdf.dg) # 1005, 87 same as cdf above, so good

>>>>

The dissolved gas data were curated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-11-10.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

@@

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2020-11-10.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) %>%
  filter(sample.source == "DG", # pull out dissolved gas and leave out gas samples
         !grepl("AK", site.id)) # omit alaska sites
```

@@

k~600~ is a standardized value that must be corrected for the differences in diffusivity among gases and water temperature at the sampling sites.  These corrections will be implemented at a later date and emission rates will be calculated using k600 for now.

@@

# although data I provided Tom (nla17gasAggregatedForPopEst2019-10-23.txt) included data from both visits,
# data used for population estimates included only visit one.  I don't know rationale for this, but will
# do the same with the revised file here.
unique(cdf$VISIT_NO) # 1
dg.1 <- dg %>% filter(visit.no == 1)

# even after filtering, dg has 85 more observations than cdf
dim(cdf) # 1005, 81
dim(dg.1) # 1090, 399

@@

dim(cdf.dg) # 1005, 86 same as cdf above, so good
